---
title: "stat431 q3 q3"
author: "Yiming Shen 20891774"
date: "18/11/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Access the NMES1988 dataset and print the first few observations
library(AER) # warnings are OK
data(NMES1988)

head(NMES1988)
help(NMES1988)
```


### (c)
```{r}
# hist for visits
hist(NMES1988$visits,
     xlab="Number of visits",
     main="The number of visits for health care",
     breaks=40,
     xlim=c(0,40),
     ylim=c(0,0.2),
     probability = T)

# Poisson Model
points(x=seq(0,40),y=dpois(x=seq(0,40),lambda = mean(NMES1988$visits)))
```

Comments: Based on the graphs, we found that the observation of 0 has much
higher density than what we expected under Poisson model. Therefore, there
are more respondents with 0 visits than might be expected under a Poisson model.


\newpage

### (d)
```{r}
# code for part (e)
require(pscl)
myZIP = zeroinfl(visits ~ chronic + health + insurance | chronic + insurance, data = NMES1988)
summary(myZIP)

```
```{R}
# chronic in Poisson part
estimate_chronic_poi <- 0.11868
exp(estimate_chronic_poi)

```


Interpretation: When number of chronic conditions increases one and holding
other variables unchanged, the expected number of physician office visits 
(rate) with current number of chronic plus one will be 1.12601 times of the 
expected number of physician office visits (rate) with current number of chronic.


```{r}
# poor health in Poisson part
estimate_poorhealth_poi <- 0.29470
exp(estimate_poorhealth_poi)
```

Interpretation: The expected number of physician office visits (rate) under
poor self-perceived health status is 1.342723 times of the expected number
of physician office visits (rate) under average self-perceived health status, 
while holding other variables unchanged.

```{r}
# intercept in logistic part
estimate_intercept_log <- -0.37426
exp(estimate_intercept_log)

```

Interpretation: When the number of chronic conditions is 0 and without private
insurance coverage, the odds of 0 physician office visit is 0.6877981, while 
holding other variables unchanged.


```{r}
# insurance in logistic part
estimate_insurance_log <- -0.88314
exp(estimate_insurance_log)

```

Interpretation: The odds ratio of 0 physician office visit with private insurance
coverage VS. without private insurance coverage is 0.4134825, while holding 
other variables unchanged.


