---
title: "stat431 a3 q4"
author: "Yiming Shen 20891774"
date: "19/11/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{R}
# step 0
N <- 1000
x1 <- rnorm(N, 0, 1)
x2 <- rbinom(N, 1, 0.25)

estimates <- matrix(NA,nrow=3, ncol=500)
CIs <- matrix(NA,nrow=6,ncol=500)

```

### (a)
```{r}
for (i in 1:500) {
  
  beta0 <- 0.5
  beta1 <- 0.2
  beta2 <- 0.5
  y <- rpois(N, exp(beta0 + beta1*x1 + beta2*x2))
  
  mod <- glm(y ~ x1 + x2, family = poisson)
  
  est_beta0 <- summary(mod)$coefficients[1,1]
  se0 <- summary(mod)$coefficients[1,2]
  
  est_beta1 <- summary(mod)$coefficients[2,1]
  se1 <- summary(mod)$coefficients[2,2]
  
  est_beta2 <- summary(mod)$coefficients[3,1]
  se2 <- summary(mod)$coefficients[3,2]
  
  estimates[1,i] <- est_beta0
  estimates[2,i] <- est_beta1
  estimates[3,i] <- est_beta2
  
  c <- 1.96
  CIs[1,i] <- est_beta0 - c*se0
  CIs[2,i] <- est_beta0 + c*se0
  CIs[3,i] <- est_beta1 - c*se1
  CIs[4,i] <- est_beta1 + c*se1
  CIs[5,i] <- est_beta2 - c*se2
  CIs[6,i] <- est_beta2 + c*se2
}

# average estimates
average_est_beta0 <- mean(estimates[1,])
average_est_beta1 <- mean(estimates[2,])
average_est_beta2 <- mean(estimates[3,])

average_est_beta0
average_est_beta1
average_est_beta2

# proportion of confidence interval coverage
coverage <- matrix(NA, nrow=3, ncol=500)

for (i in 1:500) {
  L0 <- CIs[1,i]
  U0 <- CIs[2,i]
  L1 <- CIs[3,i]
  U1 <- CIs[4,i]
  L2 <- CIs[5,i]
  U2 <- CIs[6,i]
  
  coverage[1,i] <- (beta0 >= L0) & (U0 >= beta0)
  coverage[2,i] <- (beta1 >= L1) & (U1 >= beta1)
  coverage[3,i] <- (beta2 >= L2) & (U2 >= beta2)
}

beta0_pro <- mean(coverage[1,])
beta1_pro <- mean(coverage[2,])
beta2_pro <- mean(coverage[3,])

beta0_pro
beta1_pro
beta2_pro
```


Comments: Based on the results, we found that average estimates are very close
to the true values; the proportion of confidence intervals that cover true
value is very close to 0.95 (since we constructed 95% CI).

### (b)
```{r}
N <- 1000
x1 <- rnorm(N, 0, 1)
x2 <- rbinom(N, 1, 0.25)

estimates <- matrix(NA,nrow=3, ncol=500)
CIs <- matrix(NA,nrow=6,ncol=500)


for (i in 1:500) {
  
  beta0 <- 0.5
  beta1 <- 0.2
  beta2 <- 0.5
  gamma0 <- 0.5
  gamma1 <- 0
  gamma2 <- 0.5
  
  y_star <- rpois(N, exp(beta0 + beta1*x1 + beta2*x2))
  
  pie <- exp(gamma0 + gamma1*x1 + gamma2*x2) / 
    (1 + exp(gamma0 + gamma1*x1 + gamma2*x2))
  
  w <- rbinom(N, 1, pie)
  y <- c(NA, N)
  y[w==1] <- 0
  y[w==0] <- y_star[w==0]
  
  
  # fit model
  mod <- glm(y ~ x1 + x2, family = poisson)
  
  est_beta0 <- summary(mod)$coefficients[1,1]
  se0 <- summary(mod)$coefficients[1,2]
  
  est_beta1 <- summary(mod)$coefficients[2,1]
  se1 <- summary(mod)$coefficients[2,2]
  
  est_beta2 <- summary(mod)$coefficients[3,1]
  se2 <- summary(mod)$coefficients[3,2]
  
  estimates[1,i] <- est_beta0
  estimates[2,i] <- est_beta1
  estimates[3,i] <- est_beta2
  
  c <- 1.96
  CIs[1,i] <- est_beta0 - c*se0
  CIs[2,i] <- est_beta0 + c*se0
  CIs[3,i] <- est_beta1 - c*se1
  CIs[4,i] <- est_beta1 + c*se1
  CIs[5,i] <- est_beta2 - c*se2
  CIs[6,i] <- est_beta2 + c*se2
  
}

# average estimates
average_est_beta0 <- mean(estimates[1,])
average_est_beta1 <- mean(estimates[2,])
average_est_beta2 <- mean(estimates[3,])

average_est_beta0
average_est_beta1
average_est_beta2

# proportion of confidence interval coverage
coverage <- matrix(NA, nrow=3, ncol=500)

for (i in 1:500) {
  L0 <- CIs[1,i]
  U0 <- CIs[2,i]
  L1 <- CIs[3,i]
  U1 <- CIs[4,i]
  L2 <- CIs[5,i]
  U2 <- CIs[6,i]
  
  coverage[1,i] <- (beta0 >= L0) & (U0 >= beta0)
  coverage[2,i] <- (beta1 >= L1) & (U1 >= beta1)
  coverage[3,i] <- (beta2 >= L2) & (U2 >= beta2)
}

beta0_pro <- mean(coverage[1,])
beta1_pro <- mean(coverage[2,])
beta2_pro <- mean(coverage[3,])

beta0_pro
beta1_pro
beta2_pro
```


Comments: Based on the results, we found that average estimates are not close
to the true values; the proportion of confidence intervals that cover true
value is much lower than 0.95 (we constructed 95% CI). 


### (c)
```{r}
N <- 1000
x1 <- rnorm(N, 0, 1)
x2 <- rbinom(N, 1, 0.25)

estimates <- matrix(NA,nrow=3, ncol=500)
CIs <- matrix(NA,nrow=6,ncol=500)


for (i in 1:500) {
  
  beta0 <- 0.5
  beta1 <- 0.2
  beta2 <- 0.5
  gamma0 <- 0.5
  gamma1 <- 0
  gamma2 <- 0.5
  
  y_star <- rpois(N, exp(beta0 + beta1*x1 + beta2*x2))
  
  pie <- exp(gamma0 + gamma1*x1 + gamma2*x2) / 
    (1 + exp(gamma0 + gamma1*x1 + gamma2*x2))
  
  w <- rbinom(N, 1, pie)
  y <- c(NA, N)
  y[w==1] <- 0
  y[w==0] <- y_star[w==0]
  
  
  # fit model
  require(pscl)
  mod <- zeroinfl(y ~ x1 + x2)
  
  est_beta0 <- summary(mod)$coefficients$count[1]
  se0 <- sqrt(summary(mod)$vcov[1,1])
  
  est_beta1 <- summary(mod)$coefficients$count[2]
  se1 <- sqrt(summary(mod)$vcov[2,2])
  
  est_beta2 <- summary(mod)$coefficients$count[3]
  se2 <- sqrt(summary(mod)$vcov[3,3])
  
  estimates[1,i] <- est_beta0
  estimates[2,i] <- est_beta1
  estimates[3,i] <- est_beta2
  
  c <- 1.96
  CIs[1,i] <- est_beta0 - c*se0
  CIs[2,i] <- est_beta0 + c*se0
  CIs[3,i] <- est_beta1 - c*se1
  CIs[4,i] <- est_beta1 + c*se1
  CIs[5,i] <- est_beta2 - c*se2
  CIs[6,i] <- est_beta2 + c*se2
  
}

# average estimates
average_est_beta0 <- mean(estimates[1,])
average_est_beta1 <- mean(estimates[2,])
average_est_beta2 <- mean(estimates[3,])

average_est_beta0
average_est_beta1
average_est_beta2

# proportion of confidence interval coverage
coverage <- matrix(NA, nrow=3, ncol=500)

for (i in 1:500) {
  L0 <- CIs[1,i]
  U0 <- CIs[2,i]
  L1 <- CIs[3,i]
  U1 <- CIs[4,i]
  L2 <- CIs[5,i]
  U2 <- CIs[6,i]
  
  coverage[1,i] <- (beta0 >= L0) & (U0 >= beta0)
  coverage[2,i] <- (beta1 >= L1) & (U1 >= beta1)
  coverage[3,i] <- (beta2 >= L2) & (U2 >= beta2)
}

beta0_pro <- mean(coverage[1,])
beta1_pro <- mean(coverage[2,])
beta2_pro <- mean(coverage[3,])

beta0_pro
beta1_pro
beta2_pro

```


Comments: Based on the results, we found that average estimates are very close
to the true values; the proportion of confidence intervals that cover true
value is very close to 0.95 (since we constructed 95% CI).


### (d)
```{r}
N <- 1000
x1 <- rnorm(N, 0, 1)
x2 <- rbinom(N, 1, 0.25)

estimates <- matrix(NA,nrow=3, ncol=500)
CIs <- matrix(NA,nrow=6,ncol=500)


for (i in 1:500) {
  
  beta0 <- 0.5
  beta1 <- 0.2
  beta2 <- 0.5
  gamma0 <- 0.5
  gamma1 <- 0
  gamma2 <- 0.5
  
  y_star <- rpois(N, exp(beta0 + beta1*x1 + beta2*x2))
  
  pie <- exp(gamma0 + gamma1*x1 + gamma2*x2) / 
    (1 + exp(gamma0 + gamma1*x1 + gamma2*x2))
  
  w <- rbinom(N, 1, pie)
  y <- c(NA, N)
  y[w==1] <- 0
  y[w==0] <- y_star[w==0]
  
  
  # fit model and use ad-hoc method
  mod <- glm(y ~ x1 + x2, family = poisson)
  
  estimate_dispersion <- summary(mod)$deviance / summary(mod)$df.residual
  
  est_beta0 <- summary(mod)$coefficients[1,1]
  se0_adj <- summary(mod)$coefficients[1,2] * sqrt(estimate_dispersion)
  
  est_beta1 <- summary(mod)$coefficients[2,1]
  se1_adj <- summary(mod)$coefficients[2,2] * sqrt(estimate_dispersion)
  
  est_beta2 <- summary(mod)$coefficients[3,1]
  se2_adj <- summary(mod)$coefficients[3,2] * sqrt(estimate_dispersion)
  
  estimates[1,i] <- est_beta0
  estimates[2,i] <- est_beta1
  estimates[3,i] <- est_beta2
  
  c <- 1.96
  CIs[1,i] <- est_beta0 - c*se0_adj
  CIs[2,i] <- est_beta0 + c*se0_adj
  CIs[3,i] <- est_beta1 - c*se1_adj
  CIs[4,i] <- est_beta1 + c*se1_adj
  CIs[5,i] <- est_beta2 - c*se2_adj
  CIs[6,i] <- est_beta2 + c*se2_adj
  
}

# average estimates
average_est_beta0 <- mean(estimates[1,])
average_est_beta1 <- mean(estimates[2,])
average_est_beta2 <- mean(estimates[3,])

average_est_beta0
average_est_beta1
average_est_beta2

# proportion of confidence interval coverage
coverage <- matrix(NA, nrow=3, ncol=500)

for (i in 1:500) {
  L0 <- CIs[1,i]
  U0 <- CIs[2,i]
  L1 <- CIs[3,i]
  U1 <- CIs[4,i]
  L2 <- CIs[5,i]
  U2 <- CIs[6,i]
  
  coverage[1,i] <- (beta0 >= L0) & (U0 >= beta0)
  coverage[2,i] <- (beta1 >= L1) & (U1 >= beta1)
  coverage[3,i] <- (beta2 >= L2) & (U2 >= beta2)
}

beta0_pro <- mean(coverage[1,])
beta1_pro <- mean(coverage[2,])
beta2_pro <- mean(coverage[3,])

beta0_pro
beta1_pro
beta2_pro
```

Comments: By using ad-hoc method, it does not yield valid inference.
Since the proportion of confidence intervals that cover the true value
is till far less than 0.95 after using estimated dispersion parameter.
